Reproduce anaylses of Beninca et al (2008)

### 2015-10-20 article/article.tex
update

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

@@

The original paper describes analyses of fluctuations in the abundance
of organisms in a plankton community derived from the Baltic Sea, housed
in a laboratory environment. The length of the time series (samples
every few days for 2,300 days) allowed for analyses revealing that the
observed dynamics exhibited characteristics consistent with chaos
produced by species interactions. The article concludes that stability
is not required for persistence of complex food webs, and that long-term
prediction of abundances may be fundamentally impossible. The
demonstration of chaotic dynamics and limited forecast horizons (sensu
\textcite{Petchey2015}) are important in the field of ecology, since the
ability to predict dynamics is an open question with considerable
applied importance \textcite{Petchey2015} \textcite{Mouquet2015}.

@@

@@

\begin{figure}[htbp]
\centering
\includegraphics{figures/dynamics.pdf}
\caption{\label{fig:dynamics}Observed population dynamics.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figures/unnamed-chunk-50-1.pdf}
\caption{\label{fig:divergence}Reproduced divergence rates and Lyapunov
exponents (figure 3 in the original article).}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/LE_comparison.pdf}
\caption{\label{fig:LE_comparison}Comparison of Lyapunov exponents in
the original article and this reproduction.}
\end{figure}

@@

Although we were not able to make a quantitatively accurate reproduction
of all results of the original article, the qualitative results were
largely identical. Quantitative differences may have resulted from
difference in algorithms used. For example,the original used the
\href{http://www.mpipks-dresden.mpg.de/~tisean/}{Tisean software} to
calculate Lyapunov exponents. As this was available from CRAN
\href{http://cran.r-project.org/web/packages/RTisean/index.html}{until
mid 2014} and since it is a bit less well integrated with R, we instead
use the tseriesChaos package \textcite{tseriesChaos}, which in any case
was largely inspired by the TISEAN project. There may also have been
some difference in data used for specific analyses, e.g., data with
zeros removed or not, as it was not always possible to be sure the
reproduction used exactly the same data.

>>>>

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Owen Petchey,
      Marco Plebani,
      Frank Pennekamp,
  },
            pdftitle=
{
[Re] Chaos in a long-term experiment with a plankton community
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

@@

The original paper describes analyses of fluctuations in the abundance
of organisms in a plankton community derived from the Baltic Sea, housed
in a laboratory environment. The length of the time series (samples
every few days for 2,300 days) allowed for analyses revealing that the
observed dynamics exhibited characteristics consistent with chaos
produced by non-linear species interactions. The article concludes that
stability is not required for persistence of complex food webs, and that
long-term prediction of abundances may be fundamentally impossible. The
demonstration of chaotic dynamics and limited forecast horizons (sensu
\textcite{Petchey2015}) are important in the field of ecology, since the
ability to predict dynamics is an open question with considerable
applied importance \textcite{Petchey2015} \textcite{Mouquet2015}.

@@

The original species abundance data contained errors (e.g., a few
numerica values had a comma in place of a period as the decimal
separator) that suggested that this was not the exact version of the
dataset used in the original article, or that this was the exact
dataset, but with errors corrected.

@@

\begin{figure}[htbp]
\centering
\includegraphics{figures/obs_pop_dyn.pdf}
\caption{\label{fig:dynamics}Observed population dynamics.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figures/div_rate.pdf}
\caption{\label{fig:divergence}Reproduced divergence rates and Lyapunov
exponents (figure 3 in the original article).}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/LE_comparison.pdf}
\caption{\label{fig:LE_comparison}Comparison of Lyapunov exponents,
estimated by direct method, in the original article and this
reproduction.}
\end{figure}

@@

Although we were not able to make a quantitatively accurate reproduction
of all results of the original article, the qualitative results were
largely identical. For example, all Lyapunov exponents estimated by
direct method are positive, as in the original article, consistent with
chaotic dynamics. Quantitative differences may have resulted from
difference in algorithms used. For example,the original used the
\href{http://www.mpipks-dresden.mpg.de/~tisean/}{Tisean software} to
calculate Lyapunov exponents. As this was available from CRAN
\href{http://cran.r-project.org/web/packages/RTisean/index.html}{until
mid 2014} and since it is a bit less well integrated with R, we instead
use the tseriesChaos package \textcite{tseriesChaos}, which in any case
was largely inspired by the TISEAN project. In addition, there may have
been some difference in algorithm parameters, as not all parameters
required by the functions we used were reported in the original ms.
There may also have been some difference in data used for specific
analyses, e.g., data with zeros removed or not, as it was not always
possible to be totally sure the reproduction used exactly the same data
as the original article.

In conclusion, this reproduction supports the general scientific
conclusions of the original article, but also shows how difficult can be
an accurate quantitative reproduction, even in the presence of the
extensive methodological details provided alongside the original
article.

### 2015-10-20 code/report.Rmd
update

Then plot the graphs with LE:
```{r}
LEs <- mutate(LEs, days=20, Difference=-0.5)
g1 <- ggplot(diverg, aes(x=days, y=Difference)) + geom_point() + facet_wrap(~Species) +
  geom_text(data=LEs, aes(label=round(le,3)), group=NULL)
g1
```

@@

ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
  geom_point() +
  facet_wrap(~Variable, scales="free_y" )
```

>>>>

Then plot the graphs with LE:
```{r}
LEs <- mutate(LEs, days=20, Difference=-0.5)
g1 <- ggplot(diverg, aes(x=days, y=Difference)) + geom_point() + facet_wrap(~Species) +
  geom_text(data=LEs, aes(x=60, y=-2.5, label=round(le,3)), group=NULL)
g1
```

@@

```{r}
levels(cors.long$Variable)
ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
  geom_point() +
  facet_wrap(~Variable, scales="free_y", ncol=3 )

cors.long$Variable <- factor(cors.long$Variable, sort(levels(cors.long$Variable)))
ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
  geom_point() +
  facet_wrap(~Variable, ncol=3 )

cors.long$Variable <- factor(cors.long$Variable, c("Cyclopoids", "Rotifers", "Calanoid.copepods", "Picophytoplankton",
                             "Nanophytoplankton", "Filamentous.diatoms", "Soluble.reactive.phosphorus",
                             "Total.dissolved.inorganic.nitrogen", "Bacteria", "Ostracods", "Harpacticoids",
                             "Protozoa"))
ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
  geom_point(col="red", size=3) +
  facet_wrap(~Variable, ncol=3 )
